---
title: "HW4 - CV"
author: "Tanuj Guha"
date: "2023-04-03"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

Load the data.
```{r}
library (Stat2Data)
data ("CountyHealth")
names (CountyHealth)
head (CountyHealth)
dim (CountyHealth)

CountyHealth$sqrtMDs = sqrt (CountyHealth$MDs)
```

4.8.a

```{r}
# Split the data, per the book exercise.

train.CH = CountyHealth [1:35, ]
test.CH  = CountyHealth [36:53, ]

# Set up a vector for coloring the training set black and the test set red.

train.test = c (rep (1, 35), rep (2, 18))

plot (sqrtMDs ~ Hospitals, data=CountyHealth, col=train.test)

# Fit model on the training sample

train.fit = lm (sqrtMDs ~ Hospitals, data=train.CH)
summary (train.fit)
plot (sqrtMDs ~ Hospitals, data=CountyHealth, col=train.test)
abline (train.fit)
```

Writing the equation down:

```{r}
model_equation <- function(model, ...) {
  format_args <- list(...)
  
  model_coeff <- model$coefficients
  format_args$x <- abs(model$coefficients)
  model_coeff_sign <- sign(model_coeff)
  model_coeff_prefix <- case_when(model_coeff_sign == -1 ~ " - ",
                                  model_coeff_sign == 1 ~ " + ",
                                  model_coeff_sign == 0 ~ " + ")
  model_eqn <- paste(strsplit(as.character(model$call$formula), "~")[[2]], # 'y'
                     "=",
                     paste(if_else(model_coeff[1]<0, "- ", ""),
                           do.call(format, format_args)[1],
                           paste(model_coeff_prefix[-1],
                                 do.call(format, format_args)[-1],
                                 " * ",
                                 names(model_coeff[-1]),
                                 sep = "", collapse = ""),
                           sep = ""))
  return(model_eqn)
}

model_equation(train.fit, digits = 4)
```

4.8.b

```{r}
# Predictions on the testing sample

test.predict = predict (train.fit, list (Hospitals = test.CH$Hospitals))
plot (test.CH$sqrtMDs ~ test.predict, main="Test Data with Training Fit")
abline (0, 1)

# Cross-validation correlation and R-squared

(crval.cor = cor (test.CH$sqrtMDs, test.predict))
(test.Rsq = crval.cor^2)

```

The cor() function, in this case test the correlation between the predicted data, and the actual sqrtMDs from the corresponding values in the independent variable. The R^2 value is pretty high, which indicates that the model was fit well using the training set, as seen when the model was run on the test set. 

4.8.c

```{r}
# Shrinkage

summary (train.fit)$r.squared - test.Rsq

```

In this case, the test R^2 is greater than the training R^2. Somehow, the model predicts the data better than the model was fit to the training set. However, given the magnitude of difference, I would not personally suspect any cherry picking of the test set. 

---
