---
title: "Chapter 2"
author: "Tanuj Guha"
date: "2023-02-25"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Stat2Data)
data("MetabolicRate")
head (MetabolicRate)
```
Let us create a model as asked in 2.24:

```{r}
model = lm (LogBodySize ~ LogMrate, data=MetabolicRate)
summary (model)
```
2.24 A) LogBodySize = -1.39 + 1.03 LogMrate
2.24 B) The slope parameter is significant! Generally speaking, I'd calculate t-stat, then p-val of that t-stat from DF. However, R provides 'Signif Codes', and the summary of the model inidicates that the slope parameter is significant. 
2.24 D) Model sum of squares is SSM and the Total sum of squares is SST. Since R^2 is SSM/SST, then the answer is 0.9478. The larger this value, the better the model explains the relationship between the Y and X variables. 

2.50 A) When LogBodySize is 0 then LogMrate is (139/103). Got the LogMrate value by solving for 0 = -1.39 + 1.03 LogMrate. Then, Mrate = 22.36 μW
2.50 B) Predicted interval = -1.38 (+/-) (-100.7*0.01381)
The 95% PI will be (-2.781334, 0)

[Confused about the 2.50 B answer, suspect it is incorrect: trying to understand the process from 2.4]

