---
title: "Ex 1.50"
author: "Tanuj Guha"
date: "2023-02-25"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Stat2Data)
data("CountyHealth")
head (CountyHealth)
CountyHealth$LogMD = log(CountyHealth$MDs)
```
```{r}
TestLM = lm (Hospitals ~ MDs, data=CountyHealth)
summary (TestLM)
```
We know that the number of MDs without transformation don't provide a good enough linear model. Let us see what it looks like:

```{r fig.width=7.5, fig.height=4}
par (mfrow = c(1, 2))
plot (TestLM, which=1:2)
```

Now let us create a model iwth the log transformed MDs, and check if the tranformation helped.

```{r}
TestLogLM = lm (Hospitals ~ LogMD, data=CountyHealth)
par (mfrow = c(1, 2))
plot (TestLogLM, which=1:2)
```

The log transformation is most certainly a better step: it decreases obvious outliers as indicated by the Q-Q plot meaning their distributions are now better matched. However, the Residuals/Fitted graph indicates the possibility of heteroskedasticity. 


