---
title: "Even Exercises 1.4-1.12 Long Jump"
author: "Tanuj Guha"
date: "Feb. 25, 2023"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Explore the Data

```{r}
library (Stat2Data)
data ("LongJumpOlympics2016")
head (LongJumpOlympics2016)
plot (LongJumpOlympics2016)
```

Does the distance of the Gold medal winning long jump in the Olympics depend on year?

## Fit the simple linear regression model

```{r}
ljmod = lm (Gold ~ Year, data=LongJumpOlympics2016)
summary (ljmod)
```

## Exercise 1.4

0.012508 is the slope estimate.

```{r fig.width=4.5, fig.height=3.5}
plot (Gold ~ Year, data=LongJumpOlympics2016)
abline (ljmod)
```

## Exercise 1.6
Y intercept ~ 7.4 units of Y axis. (FOR THIS GRAPH ONLY)
Note: the units in Y-axis are positive for this graph because the year 
value in X-axis doesn't begin at origin. At origin, the 'true' intercept would be -16.47


```{r}
confint (ljmod)
```

## Exercise 1.8

We expect the slope intercept to be between -21.95 and -10.98 units with 95% Confidence. 

## Exercise 1.10

 0.2595 (units of length)^2

## Exercise 1.12

26 degrees of freedom


## Assess the fit

```{r fig.width=7.5, fig.height=4}
par (mfrow = c(1, 2))
plot (ljmod, which=1:2)
```

The residuals vs fitted plot does not indicate non-linearity, and there doesn't appear to be any obvious outlier either. 
