---
title: "RStudio HW"
author: "Rick Guha"
date: "February 24, 2023"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in the data

First, we will load the data.

```{r}
# Activate the Stat2Data package (must be installed first - one-time step)
library (Stat2Data)

# load the Day1Survey data set
data (Day1Survey)

head (Day1Survey)
```

Then make a plot.

```{r}
# Three different ways to do the same thing
boxplot (Height ~ Sex, data=Day1Survey)
```

The plot above suggests that there is some difference between the groups.

Here are the summary statistics by group:

```{r}
library (mosaic)
favstats(Height ~ Sex, data=Day1Survey)
```

Let's do a two sample t-test to determine if there is actually a statistically significant difference between the heights of men and women:

```{r}
t.test (Height ~ Sex, data=Day1Survey, var.equal=T)
```
The p-value is 9.092e-10, so we reject the null hypothesis and conclude that the mean mean height difference is statistically different between men and women.

We are 95% confident that women, on average, are between -8.7 inches and -5.15 inches shorter than men.

