---
title: "Scores and Admissions"
author: "Mbonisi and Rick"
date: "`r Sys.Date()`"
output: slidy_presentation
---

## Data Aquisition
```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(xlsx)
library(datadictionary)
colleges = read_xlsx("colleges.xlsx", sheet=1)
```
This marks the end of house keeping data cleaning. 

## Data Cleaning
We know that 'NumberEnrolled’ is skewed. 
```{r}
hist(colleges$NumberEnrolled)
```

```{r}
colleges$logEnrolled <- (log(colleges$NumberEnrolled))
hist(colleges$logEnrolled)
```

## First order model
```{r}
model1 = lm(colleges$ACT ~ colleges$logEnrolled + colleges$ACTPercentage + colleges$SATWR + colleges$SATM + colleges$GPA + colleges$CollegePrep + colleges$TOEFL)
summary(model1)
```
## First Order Model Summary
a) Significant predictors: logEnrolled, ACTPercentage, SATWR and SATM.
b) There SAT scores (dis aggregated by subjects) and the percentage of students who choose to submit their ACT scores.
```{r}
numerical = cbind(ACT = colleges$ACT, logEnrolled = colleges$logEnrolled, ACTPercentage = colleges$ACTPercentage, SATM = colleges$SATM, SATWR = colleges$SATWR)
pairs(numerical)
```

## First Order Model Continued
```{r}
plot(model1)
```
Resid / Fitted Plot shows no concerning curvature.
Cook's Distance none greater than 0.5 

## Box Cox
```{r}
library(MASS)
bc = boxcox(model1)
lambda = bc$x[which.max(bc$y)]
cat(lambda)
```

## Box Cox Fix
Since Lambda = 2, therefore we run a quick fix. 
```{r}
colleges$ACT = ((colleges$ACT^lambda-1)/lambda)
model1BC = lm(colleges$ACT ~ colleges$logEnrolled + colleges$ACTPercentage + colleges$SATWR + colleges$SATM + colleges$GPA + colleges$CollegePrep + colleges$TOEFL)
summary(model1BC)
```
Significant predictors: logEnrolled, ACTPercentage, SATWR and SATM, and _now_ GPA!

## Resid Analysis on BC Fix
```{r}
plot(model1BC)
```
Fat Tailed Q-Q.

## Step-wise
```{r}
st = stepAIC(model1BC, direction="both")
```
__lm(formula = colleges$ACT ~ colleges$logEnrolled + colleges$ACTPercentage + 
    colleges$SATWR + colleges$SATM + colleges$GPA)__
    
## Stepwise Model
Create the Centered Dataset
```{r}
library(dplyr)
colleges <- colleges %>% mutate_if(is.character, as.numeric)
CollegesM = colleges %>% 
  mutate_at(.vars = colnames(colleges)[1:11], 
            .funs = list("scaled" = scale))
```

Add Interaction terms
```{r}
modelCen =lm(CollegesM$ACT ~ (CollegesM$logEnrolled_scaled + CollegesM$ACTPercentage_scaled + 
    CollegesM$SATWR_scaled + CollegesM$SATM_scaled + CollegesM$GPA_scaled)^2)
summary(modelCen)
```

## Stepwise to Interaction terms
```{r}
stInt = stepAIC(modelCen, direction="both")
summary(stInt)
```

## StepWise on Centered 
```{r}
FinalModel = lm(formula = CollegesM$ACT ~ CollegesM$logEnrolled_scaled + CollegesM$ACTPercentage_scaled + 
    CollegesM$SATWR_scaled + CollegesM$SATM_scaled + CollegesM$GPA_scaled + 
    CollegesM$logEnrolled_scaled:CollegesM$SATWR_scaled + CollegesM$logEnrolled_scaled:CollegesM$SATM_scaled + 
    CollegesM$logEnrolled_scaled:CollegesM$GPA_scaled + CollegesM$ACTPercentage_scaled:CollegesM$SATWR_scaled + 
    CollegesM$SATWR_scaled:CollegesM$SATM_scaled + CollegesM$SATWR_scaled:CollegesM$GPA_scaled + 
    CollegesM$SATM_scaled:CollegesM$GPA_scaled)
summary(FinalModel)
```

## Plots of Final Model
```{r}
plot(FinalModel)
```

