---
title: "Math 327 Chapter 9 Homework"
author: "(Your Name Here)"
date: "January 10, 2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library()
```

## Read and Plot the Data

```{r}
library (Stat2Data)
data ("Titanic")

# Remove one row with PClass = *

Titanic = Titanic [!Titanic$PClass=="*",]
head (Titanic)

plot (Titanic [, 2:5], col=Titanic$Survived*2+1)

Titanic$Died = 1 - Titanic$Survived
```

Note: It is relatively easier to explain odds ratios for positive effects (positive slopes), so the analyses in this assignment model the odds of not surviving (Died), rather than the odds of surviving.

## Question 1

```{r}
# Frequency table of Died vs Predictor
table (Titanic$Died, Titanic$Sex)
# Logistic fit
fit1 = glm (Died ~ Sex, data=Titanic, family=binomial)
summary (fit1)
# Odds Ratio for Died, Male vs. Female
exp (fit1$coefficients)
```
We know that odds ratio = e ^ (__fitted slope__).
However we can use R here to extract the odds ratio:
```{r}
exp(coef(fit1))
```


## Question 2

```{r}
# Odds Ratio confidence interval for Died, Male vs. Female
exp (confint (fit1))
```



## Question 3

```{r, fig.height=3.5, fig.width=7.5}
par (mfrow = c(1,2))
with (Titanic, boxplot (Age ~ Died, horizontal = T))
plot (jitter (Died, 0.2) ~ Age, data=Titanic)
with (Titanic [!is.na (Titanic$Age), ], lines (lowess (Died ~ Age), col='red', lty=2))
# Logistic fit
fit2 = glm (Died ~ Age, data=Titanic, family=binomial)
xseq = seq (0, 75, length=100)
lines (xseq, predict (fit2, list (Age=xseq)))
summary (fit2)
# Odds Ratio for Died
exp (fit2$coefficients)
```



## Question 4

```{r}
# Odds Ratio confidence interval for Died
exp (confint (fit2))
```



## Question 5

```{r}
# Frequency table of Died vs Predictor - Ignore the column labelled *
table (Titanic$Died, Titanic$PClass)
# Logistic fit
fit3 = glm (Died ~ PClass, data=Titanic, family=binomial)
summary (fit3)
# Odds Ratio for Died, 2nd or 3rd Class vs 1st Class
exp (fit3$coefficients)
```



## Question 6

```{r}
# Odds Ratio confidence intervals for Died
exp (confint (fit3))
```


